---
title: "Depinet Code Sample (Networks)"
author: "Jacob Depinet"
date: "2025-05-22"
output: pdf_document
---

This is an excerpt of my code from my senior data analytics capstone project (STAT 4911). Our client, Honda Motor Company, was interested in which products they were sourcing from multiple vendors and which vendors were they sourcing multiple products from. The intent was to optimize their purchasing, reducing duplicate purchases, and making their acquisition practices more efficient. I conducted network analysis and visualizations. Labels have been replaced with fake vendor/item names.

```{r}
#packages
library(dplyr)
library(tidyr)
library(ggnetwork)
```

```{r}
#load data
load(file = "C:\\Users\\jdepi\\OneDrive\\Documents\\STAT 4911\\Honda\\fixedHondaDataV2.rdata")
completeHondaDataNew <- completeHondaData
```

```{r}
#get purchasing counts
purchaseCounts <- completeHondaDataNew %>%
  group_by(vendor_id, `LEVEL 3`) %>%
  summarise(count = n()) %>%
  ungroup()

#build each node type
vendorNodes <- data.frame(id = unique(purchaseCounts$vendor_id), type = "vendor_id")
level3Nodes <- data.frame(id = unique(purchaseCounts$`LEVEL 3`), type = "LEVEL 3")
allNodes <- rbind(vendorNodes, level3Nodes)

#Get edges
edges <- purchaseCounts %>%
  mutate(from = vendor_id, to = `LEVEL 3`)

#Build network
NetworkHonda <- ggnetwork(edges, layout = "fruchtermanreingold", nodes = allNodes)

head(NetworkHonda)
```

```{r}
#Prepare Name dataset

#Vendor Nodes to highlight
#- 2000550032: 17 (Fake Vendor 1)
#- 2000550313: 11 (Fake Vendor 2)

#Item Nodes to highlight
#- Fake Item 1: 222
#- Fake Item 2: 167

relevant_node_labels <- data.frame(
  id = c("2000550032", "2000550313", "Fake Item 1", "Fake Item 2"),
  label = c("Fake Vendor 1", "Fake Vendor 2", "Fake Item 1", "Fake Item 2")
)

NetworkHonda <- merge(NetworkHonda, relevant_node_labels, by.x = "vertex.names", by.y = "id", all.x = TRUE)
```

```{r}
#build network vis (labels)
netVis <- ggplot(NetworkHonda, aes(x = x, y = y, xend = xend, yend = yend)) + 
  geom_edges(linewidth = 0.35) +
  geom_nodes(size = 0.75, aes(color = ifelse(vertex.names %in% unique(purchaseCounts$`LEVEL 3`), "LEVEL 3", "vendor_id"))) + 
  scale_color_manual(values = c("LEVEL 3" = "blue", "vendor_id" = "red")) +
  geom_nodelabel_repel(aes(label =  label)) + #I  can remove the labels
  labs(caption = "Blue: Item Description, Red: Vendor ID") +
  theme_blank(legend.position = "none")
```

```{r}
#build network vis (no labels)
netVisNoLab <- ggplot(NetworkHonda, aes(x = x, y = y, xend = xend, yend = yend)) + 
  geom_edges(linewidth = 0.35) +
  geom_nodes(size = 0.75, aes(color = ifelse(vertex.names %in% unique(purchaseCounts$`LEVEL 3`), "LEVEL 3", "vendor_id"))) + 
  scale_color_manual(values = c("LEVEL 3" = "blue", "vendor_id" = "red")) +
  labs(caption = "Blue: Item Description, Red: Vendor ID") +
  theme_blank(legend.position = "none")
```

```{r}
#visualize
knitr::knit_print(netVis)
knitr::knit_print(netVisNoLab)
```



```{r}
same_item_many_vendor <- NetworkHonda %>%
  group_by(to) %>%
  summarize(same_item_many_vendor = n_distinct(from))
```

```{r}
#Same items many vendors
ggplot(data = same_item_many_vendor %>% top_n(15,same_item_many_vendor), aes(reorder(to, same_item_many_vendor), y = same_item_many_vendor)) +
  geom_col(fill = "darkred") +
  ggtitle("Top 15 Items Shared with Number of Vendors") + 
  xlab("Item (LEVEL 3 Description)") +
  ylab("Number of Vendors Bought From") +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.y = element_text(angle = 30, size = 12), axis.title.x = element_text(size = )) +
  scale_x_discrete(labels = c("Fake Item 15", "Fake Item 14", "Fake Item 13", "Fake Item 12", "Fake Item 11", "Fake Item 10", "Fake Item 9", "Fake Item 8", "Fake Item 7", "Fake Item 6", "Fake Item 5", "Fake Item 4", "Fake Item 3", "Fake Item 2", "Fake Item 1"))
```

```{r}
same_vendor_many_items <- NetworkHonda %>%
  group_by(from) %>%
  summarize(same_vendor_many_items = n_distinct(to))
```


```{r}
#Same vendor many items
ggplot(data = same_vendor_many_items %>% top_n(15, same_vendor_many_items), aes(reorder(from, same_vendor_many_items), y = same_vendor_many_items)) +
  geom_col(fill = "darkred") +
  ggtitle("Top 18 Vendors Buying the Most Amount of Unique Items") + 
  xlab("Vendor") +
  ylab("Number of Different Items Bought by Vendor") +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.y = element_text(angle = 30)) +
  scale_x_discrete(labels = c("Fake Vendor 18", "Fake Vendor 17", "Fake Vendor 16", "Fake Vendor 15", "Fake Vendor 14", "Fake Vendor 13", "Fake Vendor 12", "Fake Vendor 11", "Fake Vendor 10", "Fake Vendor 9", "Fake Vendor 8", "Fake Vendor 7", "Fake Vendor 6", "Fake Vendor 5", "Fake Vendor 4", "Fake Vendor 3", "Fake Vendor 2", "Fake Vendor 1"))
```

```{r}
#This puts the dataframe into wide format
PreNetworkHonda <- purchaseCounts %>%
  pivot_wider(names_from = 'LEVEL 3', values_from = count, values_fill = 0)

total_products_bought <- data.frame(
  vendor_id = PreNetworkHonda[, 1],
  total_products_bought = PreNetworkHonda %>%
 reframe(total_products_bought = rowSums(select(., -1)))
)
```


```{r}
ggplot(data = total_products_bought %>% top_n(15,total_products_bought), aes( x = reorder(vendor_id, total_products_bought), y = total_products_bought)) +
  geom_col(fill = "darkred") +
  ggtitle("Top 15 Vendors Buying the Most Total Items") + 
  xlab("Vendor") +
  ylab("Number of Total Items Bought by Vendor") +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.y = element_text(angle = 30)) +
  scale_x_discrete(labels = c("Fake Vendor 15", "Fake Vendor 14", "Fake Vendor 13", "Fake Vendor 12", "Fake Vendor 11", "Fake Vendor 10", "Fake Vendor 9", "Fake Vendor 8", "Fake Vendor 7", "Fake Vendor 6", "Fake Vendor 5", "Fake Vendor 4", "Fake Vendor 3", "Fake Vendor 2", "Fake Vendor 1"))
```



























